Regularization Techniques for PSF--Matching Kernels. II. Pre--Filtering of Input Images

### 2015-02-17 submit/ms.tex
Edits from ACB third-readthrough

We describe a novel technique for reducing the numbers of false detections in difference images due to deconvolution (or sharpening) by the profile--matching kernel.
To accomplish this, we pre--smooth the input science image by its own point spread function (PSF) before performing profile--matching.
This results in a maximum--likelihood image for point source detection, which has a $\sqrt{2}$ broader point--source profile than the image PSF.
We then match a reference image to this smoothed image, effectively resulting in a point--source maximum likelihood difference image.
%The process of detection and measurement happens directly on this difference image.
By smoothing the input image, we reduce the number of conditions under which the reference image has to be sharpened to match the input image, which is a significant source of systematic false detections.
We compared this pre--filtering technique with the traditional (post--filtering) image subtraction approach using a suite of high--fidelity simulated images from the Large Synoptic Survey Telescope (LSST) PhoSim effort.
We used a configuration environment to tune each pipeline, using an objective function that minimizes the mean--square residuals in a control sample of sources in the images.
We then ran a suite of images through image subtraction and source detection and measurement, using a simulated deep template image and three simulated science images that differ primarily in seeing.
Situations that lead to a deconvolution of the template image are shown to have a lower sensitivity to true variability, and higher incidence of false detections, due to enhancement of the image noise.
In this regard, using the pre--filtering technique is clearly preferred.
Under smoothing of the template image, we find that the quality of the difference image pixels is very similar under the two approaches.
However, the propagated per--pixel variance varies significantly at the detection stage, leading to differences in the rate of false detections.
We find a single--parameter correction to the effective detection threshold that brings the run of false detections with detection threshold in--line with theoretical expectations.
We examined several sources for this discrepancy.
An over or under--estimate of the background is shown to bias the ratio of positive--going to negative--going false detections, but does not significantly effect the total numbers.
However, a misunderstanding of the image noise is shown to significantly impact the overall false detection rate.
We find that even a 2\% misestimate of the noise may change the false detection rate by a factor of 1.5--2 at the 5--sigma detection threshold.
Under both pipelines, the propagated variance in the resulting difference images is within 1\% of the empirical variance.
However, with when post--filtering the difference images for detection, the variance is underestimated by 4--5\% (when not deconvolving), leading to an increase in the rate of false detections.
This suggests that the multiple integral transforms being applied to images lead to significant misestimation of the variance, as the pixels become correlated and current algorithms do not track the pixel covariance during these operations.
Finally, we note that in order for there to be less than 1 statistical false detection per LSST difference image, the detection pipeline must operate at a threshold of 5.5--sigma under median (0.6'') seeing conditions.

@@

Real--time, wide--field surveys have rapidly become one of the main pillars of time--domain astronomy.
Multiple projects that are on--going or planned will generate terabyte to petabyte--sized databases, with their core mission statements to enable an understanding of the Universe through the study of its variability.
A real--time analysis component is critical to the success of many of these surveys, as this allows rapid recognition and followup of rare events that are only detectable due to the massive volumes of data acquired.
However, the management of this data -- the automated acquisition, processing, storage, and retrieval of it -- poses its own challenges, as the data volumes scale exponentially from one survey to the next.
Significant investments in software and algorithms are the only way to meet these challenges head--on.
For example, the Large Synoptic Survey Telescope (LSST), which was approved for construction in the Fall of 2014, has had an active Data Management group working continuously since 2004 to refine exiting algorithms and to develop new ones, to meet the challenge of first engineering light in 2020.
We report here on one of these software development projects: the study of image subtraction software for the detection of variability in astronomical images.

The detection of transient objects or known--object variability in uncrowded fields may be undertaken at the catalog level \cite{drake}.
However, when searching for variability in regions where there is significant confusion, catalog--based methods become heavily dependent on how the complex objects are deblended, which itself is typically a function of the image resolution (or seeing).
For this reason, image subtraction techniques have been developed that perform the subtraction of a prior imaging epoch from a newly acquired one, subtracting out all static sources and only leaving signatures of objects that have changed in position or brightness in a ``difference image''.
By reducing the number of objects to measure, these variable sources may be uniquely photometered without significant deblending issues \citep[however, crowded difference images with spatialy overlapping variable sources are possible, e.g.][]{m31}.
This technique is highly advantageous for the study of variability in the Galactic bulge \citep[e.g. microlensing;][]{microlensing}, nearby galaxies \citep[e.g. nearby supernovae;][]{nearby-sne}, or for the study of nuclear variability in more distant galaxies \cite[e.g.][]{yumi}.
When used, image subtraction is typically applied to the entirety a time--domain dataset, in both sparse and crowded fields, so as to have a consistent treatment of all of the data.

It is both an opportunity and a challenge that image subtraction reveals everything that has varied between two images, since that also includes the atmosphere, the telescope pointing, the temperature of the detectors, and the condition of the detectors themselves.
This dependency makes it absolutely necessary to calibrate the science pixels, a requirement that becomes more stringent as data volumes become larger and rare effects become commonplace.
Assuming that the instrumental effects are understood, the core challenge to image subtraction is to match the point--spread--functions (PSFs) of the two images being compared.
There are multiple ways to perform PSF--matching, and techniques that operate in the spatial domain \citep{alard-lupton,alard} have largely been preferred over techniques that operate in the Fourier domain \citep{tomaney-crotts}.
The application of algorithms to the images may also cause artificial systematic sources of variability, including misapplication of the image subtraction software itself.
This is clearly a circumstance to be avoided, and the motivation behind this work.

A key recognition of recent surveys is that systematic effects (i.e. non--astrophysical variability) dominate the population of objects found in difference images \citep{sdss2,panstarrs}.
This has opened the door for the adoption of machine--learning (ML) techniques to separate the true transients from the false detections\footnote{In binary classification, sources of contamination are referred to as ``false positives''.  However, since ``positive'' also may refer to the polarity of a difference image detection, and we indeed accept both positive--going and negative--going detections as ``false positives'', we use the term ``false detections'' here to refer to these sources.}\citep{josh}.
This work is encouraging, with the most recent ML suites having XXX completeness with YYY contamination.
However, for LSST this level of judgment must be improved, with the SRD requiring XXX and YYY for the nightly real--time alert stream.
A necessary step towards achieving this goal is to understand the sources of false detections in images, and to develop new techniques to avoid generating or reporting them.
We describe such a novel algorithm here, which involves regularizing the PSF--matching operation by smoothing the science image with its own PSF.
As we will demonstrate, this is not an arbitrary decision, but merely re--orders an operation that already occurs naturally in the process of difference image source detection.
By moving this operation earlier in the pipeline in a manner that commutes with the standard technique, we introduce no new pixel--level operations into the processing, while simultaneously reducing our sensitivity to systematic effects due to ``sharpening'', or deconvolution, during PSF--matching.

@@

Image subtraction proceeds by assuming that science image $S(x,y)$ can be modeled as a convolution of a reference template image $T(x,y)$ by a PSF--matching kernel $K(u,v;x,y)$ (indices $u,v$ indicate that the kernel itself is a 2--dimensional function, which varies as a function of position $x,y$ in the image.
During convolution and correlation there is an implicit summation over $u,v$).
The two images will have different point--spread--functions (PSFs), which are the time--averaged transfer functions of a point source through the Earth's atmosphere, telescope optics, and into the silicon of the detector before being read out.
The essence of image subtraction is to match the PSFs of these two images using a $K(u,v;x,y)$ so that they may be subtracted pixel by pixel.

In practice, it is assumed that the PSF--matching kernel may be decomposed using a set of basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the coefficients in front of each basis are determined through:
\begin{eqnarray}
\label{eq-soln}
C_i & \equiv & (K_i \otimes T) \\ 
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\
b_{i}  & = & M_{ij} a_{j}. \nonumber
\end{eqnarray}
In these equations, the terms $C_i$ come from a convolution of the template image with each basis function, and $\sigma^2(x,y)$ are the per--pixel variances.
The solution for $K(u,v)$ is determined by solving for each coefficient $a_i$, which represent the relative weights of each basis.

Point--spread--functions typically have spatial variation across an image due to high--order optical distortions and structure in the perturbing atmosphere; hence the PSF--matching kernel must also vary spatially.
To generate this model, it is assumed that the weights of the basis functions $a_i$ themselves vary spatially, i.e. $K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$.
In general, there is also a differential background map between the two images $b(x,y)$ that may be fit for using a low--order polynomial.
The final image difference is then calculated through $D(x,y) = S(x,y) - T(x,y) \otimes K(u,v;x,y) - b(x,y)$.

It is important to note that the choice of basis functions $K_i(u,v)$ is a degree of freedom in this problem.
Many implementations \citep{Alard98,Alard00} use a set of 3 Gaussians, each with a different width $\sigma_n$, and each modified by a Laguerre polynomial to a given order.
Subsequent studies \citep[e.g.][]{2007AN....328...16I} have suggested that a constant ratio be maintained between the different Gaussian widths, such that $\sigma_{n+1} = \beta \times \sigma_{n}$.
We adopted the value $\beta = 2.0$ for this work.
We set the overall scale for these widths by noting that, under the assumption that the PSFs of the images are themselves Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the template image), the width of the matching kernel $\sigma_K$ should be simply $\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.
We used this width for the central Gaussian in a 3--component basis, with the smaller and larger Gaussians a factor of $1/\beta$ and $\beta$ different, respectively.

Detection on a difference image occurs after correlation of $D(x,y)$ with the science image's PSF, yielding a detection image $D'(x,y) = D(x,y) \circ PSF_S(u,v;x,y)$, where the $\circ$ operator denotes correlation.
The values of the pixels in $D'(x,y)$ provide a maximum--likelihood estimate of there being a point source at that position.
Detection occurs by searching for single pixels in $D'(x,y)$ that are more than $N$ times the square root of the per--pixel background variance.
The value of $N$ defines the detection threshold, and is also a degree of freedom in the problem, chosen to minimize the number of statistical fluctuations while maximizing the sensitivity for true sources of variability.
Once sources are detected in image $D'(x,y)$, measurement proceeds on the associated pixels in $D(x,y)$.

In this work we investigated two orders of operations for the optimal point--source filtering by the PSF.
In the first method, which is the classical implementation of the technique, we create a difference image that is then post--filtered with the PSF for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K_{Post}(u,v;x,y)  \nonumber \\
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ PSF_S(u,v;x,y).  \nonumber
\end{eqnarray}
In the second method, which we refer to as pre--filtering, we match the template image to a PSF--filtered science image:
\begin{eqnarray}
D_{Pre}(x,y)  & = & S(x,y) \circ PSF_S(u,v;x,y) - \nonumber \\
             &   & T(x,y) \otimes K_{Pre}(u,v;x,y) \nonumber \\
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber
\end{eqnarray}
In this case, both detection and measurement happen on the same image.
For model--based measurements, the model will have to be smoothed with the image PSF before being compared to the pixels.

@@

In order to test the efficacy of these two approaches, we generated simulated images using the LSST PhoSim package v3.2.5 \citep{phosim}.
These images are high--fidelity simulations of the end--to--end LSST system, including effects from the atmosphere, optics, and detectors.
In order to establish an algorithmic baseline for the image subtraction code, we turned off certain confounding elements of the simulations, including airglow variations, cloud screens, tracking variations, quantum efficiency variations, diffraction, saturation, and blooming.
An input stellar catalog was generated to seed the simulations that included a random field with $\sim 1000$ stars for each 4000x4072 pixel LSST sensor, and a flat $r$-band magnitude range of $19<r<21$.
All stars in the sample had identical spectral energy distributions to avoid chromatic effects.
% km50_5000.fits_g20_5140.gz
For all sources, proper motion, parallax, and variability were set to zero.
Thus there is no intrinsic variability of any sort designed into the simulated images, and any detections in their differences will be by--definition false detections.

@@

A given image consists of three different layers, or planes, that track different bits of information.
The first plane contains the science pixels, which is the layer on which detection and measurement algorithms operate.
The second plane consists of a pixel--wise bitmask, indicating the quality of each pixel (was it saturated, or a detector defect) as well as detection information (was this pixel part of a source).
The third plane contains the pixel--wise variance, which is tracked through all additive, multiplicative, and integral transform operations.
The ``empirical'' variance of an image is estimated by taking all of the unmasked pixels and calculating their interquartile range, scaling by 0.741 (appropriate for a Gaussian distribution) and squaring the result.
The ``propagated'' variance of an image is estimated by taking all of the unmasked variance pixels, and calculating their mean or median.

@@

At the start of processing, there are a pair of template and science images that are to be astrometrically and photometrically registered.
The process of astrometric registration refers to the exact alignment of the two images in pixel space, while photometric alignment refers to matching of the PSF shape, sky backgrounds, and zero--point.
From the input source catalog we choose 80\% of the stars to use in the registration process, with the remaining 20\% of the objects serving as a control sample to assess the effectiveness of $K(u,v;x,y)$ to subtract objects that were {\it not} used in the modeling process.
A direct image--to--image relative astrometric mapping was determined using the 80\% sample, and a sinc--based resampling kernel applied to the template image.
The resulting warped image was used as the image subtraction template $T(x,y)$.
The science image $S(x,y)$ and $T(x,y)$ were sent to the image subtraction task, whose purpose is to fit for $K(u,v;x,y)$ and to produce the difference image $D(x,y)$.

The first step in this process was to use the image PSF widths to define the {\it sizes} of the Gaussians' shapes in $K_i(u,v)$.
It is important to note that these widths are not fit parameters that are optimized under Equation~\ref{eq-soln}, and must be chosen by the user beforehand.
When $\sigma_S > \sigma_T$, a three--Gaussian basis was chosen, with the central Gaussian width being $\sqrt{\sigma_S^2 - \sigma_T^2}$, and smaller and larger Gaussians that were scaled in width by a factor of $\beta = 2$.
The degree of the modifying Laguerre polynomials were set using the mean--squared--error analysis below, with the smallest Gaussian typically modified by a degree 4--6 polynomial, and the larger Gaussians modified to lower order.
%We modified each of the Gaussians by a set of Laguerre polynomials of a given order.
%The smallest Gaussian was modified to the specified order, with the others modified by floor(order/2).
The total number of basis shapes in a given set is $\sum_{n=1}^{3} ({\rm degree}_n+1)\times({\rm degree}_n+2)/2$.

@@

The dimensions of the PSF matching kernel were chosen to be 6 times the largest Gaussian width, with a minimum kernel size of 21x21 pixels.
For kernels that are significantly smaller than this, the Gaussians have significant (non--zero) power at the kernel boundaries, leading to square systematic artifacts at the scale of the kernel in the difference images.

@@

The ensemble of sources was then used to constrain a spatial model of the kernel $K(u,v;x,y)$.
For the spatial kernel model, we assumed that each of the kernel coefficients $a_i$ may be represented by a $N^{th}$ order 2--dimensional Chebyshev polynomial.
Having generated the full spatial solution $K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$, we evaluated the spatial kernel solution at the position of each source, created a new difference image using this kernel, and recalculated the metrics defined above.
These are referred to as the {\tt SPATIAL} kernel metrics.
Importantly, we also interpolated this solution to the positions of the control sample, which were explicitly {\it not} used in the spatial fit, and evaluated these same metrics.

The differences between the {\tt LOCAL} and {\tt SPATIAL} metrics for the training sample, as well as the values of the {\tt SPATIAL} metrics for the control sample, reflect the appropriateness of the spatial model and our ability to interpolate and extrapolate to the full extent of the images.
We used the mean square error (MSE) of the control sample to examine the tradeoff between bias and variance in the spatial interpolation.
We define the bias as $\left| data - model \right|$, the variance as $(data - model)^2$, with the MSE as {\tt bias$^2$ + variance}.  
In this context, the bias is the mean of the difference image, and the variance is the mean square of the difference image.
We found that N = 5 or 6 was sufficient to minimize the control sample MSE, or maximize the predictive power of the spatial model without overfitting.

Figure~\ref{fig:1} presents four sets of common image subtraction failure modes, and associated choices in pipeline configuration that may cause them.

>>>>

We describe a novel technique for reducing the numbers of false detections in difference images due to deconvolution (or sharpening) by the profile--matching kernel.
To accomplish this, we pre--smooth the input science image by its own point spread function (PSF) before performing profile--matching.
This results in a maximum--likelihood image for point source detection, which has a $\sqrt{2}$ broader point--source profile than the image PSF.
We then match a reference image to this smoothed image, effectively resulting in a point--source maximum likelihood difference image.
%The process of detection and measurement happens directly on this difference image.
By smoothing the input image, we reduce the number of conditions under which the reference image has to be sharpened to match the input image, which is a significant source of systematic false detections.
We compared this pre--filtering technique with the traditional (post--filtering) image subtraction approach using a suite of high--fidelity simulated images from the Large Synoptic Survey Telescope (LSST) PhoSim effort.
We used a configuration environment to tune each pipeline, using an objective function that represents the mean--square residuals in a control sample of sources in the images.
We then ran a suite of images through image subtraction and source detection and measurement, using a simulated deep template image and three simulated science images that differ primarily in seeing.
Situations that led to a deconvolution of the template image had a lower sensitivity to true variability, and higher incidence of false detections, due to enhancement of the image noise.
In this regard, using the pre--filtering technique is clearly preferred.
Under smoothing of the template image, we find that the quality of the difference image pixels is very similar under the two approaches.
However, the propagated per--pixel variance varies significantly at the detection stage, leading to differences in the rate of false detections.
We find a single--parameter correction to the effective detection threshold that brings the run of false detections with detection threshold in--line with theoretical expectations.
We examined several sources for this discrepancy.
An over or under--estimate of the background is shown to bias the ratio of positive--going to negative--going false detections, but does not significantly affect the total numbers.
However, misestimates of the image noise are shown to significantly impact the overall false detection rate.
We find that even a 2\% misestimate of the noise may change the false detection rate by a factor of 1.5--2 at the 5--sigma detection threshold.
Under both pipelines, the propagated variance in the resulting difference images is within 1\% of the empirical variance.
However, with when post--filtering the difference images for detection, the variance is underestimated by 4--5\% (when not deconvolving), leading to an increase in the rate of false detections.
This suggests that the multiple integral transforms being applied to images lead to significant misestimation of the variance, as the pixels become correlated and current algorithms do not track the pixel covariance.
Finally, we note that in order for there to be less than 1 statistical false detection per LSST difference image, the detection pipeline must operate at a threshold of 5.5--sigma under median (0.6'') seeing conditions.

@@

Real--time, wide--field surveys have rapidly become one of the main pillars of time--domain astronomy.
Multiple projects that are on--going or planned will generate terabyte to petabyte--sized databases, with core objectives that include an understanding of the Universe through the study of its variability.
A real--time analysis component is critical to the success of many of these surveys, as this allows rapid recognition and followup of rare events that are only detectable due to the massive volumes of data acquired.
However, the management of this data -- the automated acquisition, processing, storage, and retrieval of it -- poses its own challenges, as the data volumes scale exponentially from one survey to the next.
Significant investments in software and algorithms are the only way to meet these challenges head--on.
For this reason, the Large Synoptic Survey Telescope (LSST), which was approved for construction in the Fall of 2014, has had an active Data Management group working continuously since 2004 to refine exiting algorithms and to develop new ones, to meet the challenge of engineering first light in 2020.
We report here on one of these software development projects: the study of image subtraction software for the detection of variability in astronomical images.

The detection of transient objects or known--object variability in uncrowded fields may be undertaken at the catalog level \cite{drake}.
However, when searching for variability in regions where there is significant confusion, catalog--based methods become heavily dependent on how objects are deblended, which itself is typically a function of the image resolution (or seeing).
For this reason, image subtraction techniques have been developed that perform the subtraction of a prior imaging epoch from a newly acquired one, subtracting out all static sources and only leaving signatures of objects that have changed in position or brightness in a ``difference image''.
By reducing the number of objects to measure, these variable sources may be uniquely photometered without significant deblending issues \citep[however, crowded difference images with spatially overlapping variable sources are possible, e.g.][]{m31}.
This technique is highly advantageous for the study of variability in the Galactic bulge \citep[e.g. microlensing;][]{microlensing}, nearby galaxies \citep[e.g. nearby supernovae;][]{nearby-sne}, or for the study of nuclear variability in more distant galaxies \cite[e.g.][]{yumi}.
When used, image subtraction is typically applied to the entirety a time--domain dataset, in both sparse and crowded fields, so as to have a consistent treatment of all of the data.

It is both an opportunity and a challenge that image subtraction reveals everything that has varied between two images, since that also includes the atmosphere, the telescope pointing, the temperature of the detectors, and the condition of the detectors themselves.
This dependency makes it absolutely necessary to calibrate the science pixels, a requirement that becomes more stringent as data volumes become larger and rare effects become commonplace.
Assuming that the instrumental effects are understood, the core challenge to image subtraction is to match the point--spread--functions (PSFs) of the two images being compared.
There are multiple ways to perform PSF--matching, but techniques that operate in the spatial domain \citep{alard-lupton,alard} have largely been preferred over techniques that operate in the Fourier domain \citep{tomaney-crotts}.
The application of algorithms to the images may also cause artificial systematic sources of variability, including misapplication of the image subtraction software itself.
This is clearly a circumstance to be avoided, and the motivation behind this work.

A key recognition of recent surveys is that systematic effects (i.e. non--astrophysical variability) dominate the population of objects found in difference images \citep{sdss2,panstarrs}.
This has opened the door for the at--scale adoption of machine--learning (ML) techniques to separate the true transients from the false detections\footnote{In binary classification, sources of contamination are referred to as ``false positives''.  However, since ``positive'' also may refer to the polarity of a difference image detection, and we indeed accept both positive--going and negative--going detections as ``false positives'', we use the term ``false detections'' here to refer to these sources.}\citep{josh}.
This work is encouraging, with the most recent ML suites having XXX completeness with YYY contamination \citep{ml}.
However, for LSST this level of judgment must be improved, to meet the requirements of XXX and YYY for the nightly real--time alert stream \citep{srd}.
A necessary step towards achieving this goal is to understand the sources of false detections in images, and to develop new techniques to avoid generating or reporting them.
We describe such a novel algorithm here, which involves regularizing the PSF--matching operation by smoothing the science image with its own PSF.
As we will demonstrate, this is not an arbitrary decision, but merely re--orders an operation that occurs naturally in the process of difference image source detection.
By moving this operation earlier in the pipeline in a manner that commutes with the standard technique, we introduce no new pixel--level operations into the processing, while simultaneously reducing our sensitivity to systematic effects due to ``sharpening'', or deconvolution, during PSF--matching.

@@

Image subtraction proceeds by assuming that science image $S(x,y)$ can be modeled as a convolution of a reference template image $T(x,y)$ by a PSF--matching kernel $K(u,v;x,y)$ (indices $u,v$ indicate that the kernel itself is a 2--dimensional function, which varies as a function of position $x,y$ in the image.
During convolution and correlation there is an implicit summation over $u,v$).
The two images will have different point--spread--functions, which are the time--averaged transfer functions of a point source through the Earth's atmosphere, telescope optics, and into the silicon of the detector before being read out.
The essence of image subtraction is to match the PSFs of these two images by fitting for a $K(u,v;x,y)$, so that they may be subtracted pixel by pixel.

In practice, it is commonly assumed that the PSF--matching kernel may be decomposed using a set of basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the coefficients in front of each basis are determined through:
\begin{eqnarray}
\label{eq-soln}
C_i & \equiv & (K_i \otimes T) \\ 
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\
b_{i}  & = & M_{ij} a_{j}. \nonumber
\end{eqnarray}
In these equations, the terms $C_i$ come from a convolution of the template image with each basis function, and $\sigma^2(x,y)$ are the per--pixel variances.
The solution for $K(u,v)$ is determined by solving for each coefficient $a_i$, which represent the relative weights of each basis.

Point--spread--functions typically have spatial variation across an image due to high--order optical distortions and structure in the perturbing atmosphere; hence the PSF--matching kernel must also vary spatially.
To generate this model, it is assumed that the weights of the basis functions $a_i$ themselves vary spatially, i.e. $K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$.
In general, there is also a differential background map between the two images $b(x,y)$ that may be fit for using a low--order function.
The final image difference is then calculated through $D(x,y) = S(x,y) - T(x,y) \otimes K(u,v;x,y) - b(x,y)$.

It is important to note that the choice of basis functions $K_i(u,v)$ is a degree of freedom in this problem.
Many implementations \citep{Alard98,Alard00} use a set of 3 Gaussians, each with a different width $\sigma_n$, and each modified by a Laguerre polynomial to a given order.
Other basis functions are indeed possible, including the most flexible option of having each pixel in the kernel be a separate basis \citep{bramich,becker}.
For the sum--of--Gaussian basis, \cite{2007AN....328...16I} have suggested that a constant ratio be maintained between the different Gaussian widths, such that $\sigma_{n+1} = \beta \times \sigma_{n}$.
We adopted the value $\beta = 2.0$ for this work.
We set the overall scale for these widths by noting that, under the assumption that the PSFs of the images are themselves Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the template image), the width of a pure--Gaussian matching kernel $\sigma_K$ should be simply $\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.
We used this $\sigma_K$ for the central Gaussian in a 3--component basis, with the smaller and larger Gaussians a factor of $1/\beta$ and $\beta$ different, respectively.

Point--source detection on a difference image occurs after correlation of $D(x,y)$ with the science image's PSF, yielding a detection image $D'(x,y) = D(x,y) \circ PSF_S(u,v;x,y)$, where the $\circ$ operator denotes correlation.
The values of the pixels in $D'(x,y)$ provide a maximum--likelihood estimate of there being a point source at that position.
Detection occurs by searching for single pixels in $D'(x,y)$ that are more than $N$ times the square root of the per--pixel background variance.
The value of $N$ defines the detection threshold, and is also a degree of freedom in the problem, chosen to minimize the number of statistical fluctuations while maximizing the sensitivity for true sources of variability.
Once sources are detected in image $D'(x,y)$, measurement proceeds on the associated pixels in $D(x,y)$.

In this work we investigated two orders of operations for the optimal point--source filtering by the PSF.
In the first method, which is the classical implementation of the technique, we create a difference image that is then post--filtered with the PSF for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K_{Post}(u,v;x,y)  \nonumber \\
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ PSF_S(u,v;x,y).  \nonumber
\end{eqnarray}
In the second method, which we refer to as pre--filtering, we match the template image to a PSF--filtered science image:
\begin{eqnarray}
D_{Pre}(x,y)  & = & S(x,y) \circ PSF_S(u,v;x,y) - \nonumber \\
             &   & T(x,y) \otimes K_{Pre}(u,v;x,y) \nonumber \\
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber
\end{eqnarray}
In this case, both detection and measurement happen on the same image.
For model--based measurements, the models will have to be smoothed with the image PSF before being compared to the pixels in $D'(x,y)$.

@@

In order to test the efficacy of these two approaches, we generated simulated images using the LSST PhoSim package v3.2.5 \citep{phosim}.
These images are high--fidelity simulations of the end--to--end LSST system, including effects from the atmosphere, telescope optics, and photon detectors.
In order to establish an algorithmic baseline for the image subtraction code, we turned off certain confounding elements of the simulations, including airglow variations, cloud screens, tracking variations, quantum efficiency variations, diffraction, saturation, and blooming.
An input stellar catalog was generated to seed the simulations that included a random field with $\sim 1000$ stars for each 4000x4072 pixel LSST sensor, and a flat $r$-band magnitude range of $19<r<21$.
All stars in the sample had identical spectral energy distributions to avoid chromatic effects.
% km50_5000.fits_g20_5140.gz
For all sources, proper motion, parallax, and variability were set to zero.
Thus there is no intrinsic variability of any sort designed into the simulated images, and any detections in their differences will be by--definition false detections.

@@

A given image consists of three different layers, or planes, that track different units of information.
The first plane contains the science pixels, which is the layer on which detection and measurement algorithms operate.
The second plane consists of a pixel--wise bitmask, indicating the quality of each pixel (was it saturated, or a detector defect) as well as detection information (was this pixel part of a source).
The third plane contains the pixel--wise variance, which is tracked through all additive, multiplicative, and integral transform operations.
The ``empirical'' variance of an image is estimated by taking all of the unmasked image--plane pixels and calculating their interquartile range, scaling by 0.741 (appropriate for a Gaussian distribution) and squaring the result.
The ``propagated'' variance of an image is estimated by taking all of the unmasked variance--plane pixels, and calculating their mean or median.

@@

At the start of processing, there are a pair of template and science images that are to be astrometrically and photometrically registered.
The process of astrometric registration refers to the exact alignment of the two images in pixel space, while photometric alignment refers to matching of the PSF shape, sky backgrounds, and zero--point.
From the input source catalog we choose 80\% of the stars to use in the registration process, with the remaining 20\% of the objects serving as a control sample to assess the effectiveness of $K(u,v;x,y)$ to subtract objects that were {\it not} used in the modeling process.
A direct image--to--image relative astrometric mapping was determined using the 80\% sample, and a sinc--based resampling kernel applied to the template image to align it with the science image at the pixel level.
The resulting warped image was used as the image subtraction template $T(x,y)$.
The science image $S(x,y)$ and $T(x,y)$ were sent to the image subtraction task, whose purpose is to fit for $K(u,v;x,y)$ and to produce the difference image $D(x,y)$.

The first step in this process was to use the image PSF widths to define the {\it sizes} of the Gaussians' shapes in $K_i(u,v)$.
It is important to note that these widths are not fit parameters that are optimized under Equation~\ref{eq-soln}, and must be chosen by the user beforehand.
When $\sigma_S > \sigma_T$, a three--Gaussian basis was chosen, with the central Gaussian width being $\sqrt{\sigma_S^2 - \sigma_T^2}$, and smaller and larger Gaussians that were scaled in width by a factor of $\beta = 2$.
The degree of the modifying Laguerre polynomials were set using the mean--squared--error analysis described below, with the smallest Gaussian typically modified by a degree 4--6 polynomial, and the larger Gaussians modified to lower order.
%We modified each of the Gaussians by a set of Laguerre polynomials of a given order.
%The smallest Gaussian was modified to the specified order, with the others modified by floor(order/2).
The total number of basis shapes in a given set is $\sum_{n=1}^{3} ({\rm degree}_n+1)\times({\rm degree}_n+2)/2$.

@@

The dimensions of the PSF matching kernel were chosen to be 6 times the largest Gaussian sigma, with a minimum kernel size of 21x21 pixels.
For kernels that are significantly smaller than this, the Gaussians have significant (non--zero) power at the kernel boundaries, leading to square systematic artifacts at the scale of the kernel in the difference images.

@@

The ensemble of sources was then used to constrain a spatial model of the kernel $K(u,v;x,y)$.
For the spatial kernel model, we assumed that each of the kernel coefficients $a_i(x,y)$ may be represented by a $N^{th}$ order 2--dimensional Chebyshev polynomial.
Having generated the full spatial solution $K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$, we evaluated the spatial kernel solution at the position of each source, created a new difference image using this kernel, and recalculated the metrics defined above.
These are referred to as the {\tt SPATIAL} kernel metrics.
Importantly, we also interpolated this solution to the positions of the control sample, which were explicitly {\it not} used in the spatial fit, and evaluated these same metrics.

The differences between the {\tt LOCAL} and {\tt SPATIAL} metrics for the training sample, as well as the values of the {\tt SPATIAL} metrics for the control sample, reflect the appropriateness of the spatial model and our ability to interpolate and extrapolate it to arbitrary regions of the image.
We used the mean square error (MSE) of the control sample to examine the tradeoff between bias and variance in the spatial interpolation.
We define the bias as $\left| data - model \right|$, the variance as $(data - model)^2$, with the MSE as {\tt bias$^2$ + variance}.  
In this context, the bias is the mean of the difference image, and the variance is the mean square of the difference image.
We found that N = 5 or 6 was sufficient to minimize the control sample MSE, or to maximize the predictive power of the spatial model without overfitting.

Figure~\ref{fig:1} presents four sets of common image subtraction failure modes we encountered, and associated choices in pipeline configuration that may cause them.

\begin{figure*}[!ht]
  \includegraphics[width=0.5\textwidth, height=0.25\textwidth]{fig1a_crop.eps}
  \includegraphics[width=0.5\textwidth, height=0.25\textwidth]{fig1b_crop.eps} \\
  \includegraphics[width=0.5\textwidth, height=0.25\textwidth]{fig1c_crop.eps}
  \includegraphics[width=0.5\textwidth, height=0.25\textwidth]{fig1d_crop.eps} \\
\caption{Four panels showing example failure modes of the image subtraction software.
  Each panel shows a pair of difference images; 
  in the {\it left} subpanel is the residual from the {\tt LOCAL} fitting, and in the {\it right} the residual from the {\tt SPATIAL} fitting.
  Starting in the upper left and proceeding clockwise, the first panel shows the residuals when deconvolving the template image; 
  the ringing in both solutions is characteristic of the deconvolution process.
  Upper right: a set of difference images where the kernel dimensions (kernel size) are too small given the widths of the Gaussians, such that the value of the kernel does not fall to zero along its boundary.
  This leaves square kernel--sized residuals around each object.
  Lower right: a set of images where the kernel shape (Gaussian sigma) is inappropriate to match the sources; 
  note that the {\tt LOCAL} residuals also show circular residuals, indicating the basis set itself is at fault (as opposed to the spatial model).
  Lower left: a set of images that demonstrate how the residuals degrade when the spatial kernel model is inappropriate.
  Note that the {\tt LOCAL} residuals appear to be white noise, while the {\tt SPATIAL} residuals show a clear dipole signature.
  For this reason, a comparison between the {\tt LOCAL} and {\tt SPATIAL} residuals is a useful diagnostic of the spatial model.
  The dipoles may also arise due to problems in the astrometric registration, or for individual extreme--color objects objects due to differential chromatic refraction.
}
\label{fig:1}
\end{figure*}
